Sperm should evolve to make female meiosis fair.

### 2014-07-16 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
graham's changes to intro

It is therefore somewhat surprising that despite the intense evolutionary pressure on female meiosis to prevent meiotic drive, 
it is potentially open to sabotage by a virtual stranger -- a haploid sperm genome.
That is, in many animal species, female meiosis is completed only after fertilization \citep{Masui_book}, 
	creating ample opportunity for interaction between the sperm and
	female meiotic machinery (note that the alternation of generations
	likely precludes this interaction in plants). 
\gc{It is becoming
        increasingly clear that sperm bring a wide variety of RNA and proteins into the
        egg REF. Some of these have known functions, for example, in most
        animal species sperm-- not eggs-- are
        responsible for the transmission of the centrosome, a vitial
        component of the mitotic machinary for the zygote. However,
        the function of many of th products the sperm beings into the
        egg is completely unknown and 
      these products vary widely over species. 
The existance of paternal effect mutations in model systems highlights
 the fact that some of these components have a wide-range of functions in
 the completion of synagamy and the development of the zygote. 
Thus sperm could potentially have functional effects on
various aspects of female meiosis, e.g. in some systems 
sperm actually carry the factor responsibly for the resumption of
female meiosis, rather than this being a simple consequence of egg
activation. } 
Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females %female meiosis %(a  heteromorphic dyad), 
     could presumably rapidly spread through a population. 
At first sight, female meiosis appears primed for exploitation by such selfish systems. Here we ask if sperm do indeed evolve to collaborate with female drivers to exploit this apparent weakness in the defense against meiotic drive. \newline 
%Why then is the requirement of fertilization to complete female meiosis so ubiquitous?

>>>>

It is therefore somewhat surprising that despite the intense evolutionary pressure on female meiosis to prevent meiotic drive, 
it is potentially open to sabotage by a virtual stranger -- a haploid sperm genome.
That is, in many animal species, female meiosis is completed only after fertilization \citep{Masui_book}, 
	creating ample opportunity for interaction between the sperm and
	female meiotic machinery (note that the alternation of generations
	likely precludes this interaction in plants). 
\gc{It is becoming
        increasingly clear that sperm bring a wide variety of RNA and proteins into the
        egg REF. Some of these have known functions, for example, in most
        animal species sperm-- not eggs-- are
        responsible for the transmission of the centrosome, a vitial
        component of the mitotic machinary for the zygote. However,
        the function of many of the products the sperm brings into the
        egg is completely unknown and 
      these products vary widely over species. 
The existance of paternal effect mutations in model systems highlights
 the fact that some of these components have a wide-range of functions in
 the completion of synagamy and the development of the zygote. 
Thus sperm could potentially have functional effects on
various aspects of female meiosis, e.g. in some systems 
sperm actually carry the factor responsibly for the resumption of
female meiosis, rather than this being a simple consequence of egg
activation. } 
Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females %female meiosis %(a  heteromorphic dyad), 
     could presumably rapidly spread through a population. 
At first sight, female meiosis appears primed for exploitation by such selfish systems. Here we ask if sperm do indeed evolve to collaborate with female drivers to exploit this apparent weakness in the defense against meiotic drive. \newline 
%Why then is the requirement of fertilization to complete female meiosis so ubiquitous? 
\gc{THERE may be too much bio. detail in previous paragraph}
